### 2023-08-22 mortality/analysis.Rmd
update

---
title: "mortality_analysis"
author: "Bohan Zhang"
date: "2023-08-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
# ts and error representator
source("../R/construct_hierarchy.R", chdir = TRUE)

store_all <- NULL
for (batch in 0:11) {
  store_all <- readRDS(paste0("store_", batch, ".rds"))$output %>%
    output_pre() %>% mutate(batch = batch) %>%
    rowwise() %>%
    mutate(bottom_mean = mean(bottom)) %>%
    ungroup() %>%
    rbind(store_all)
}
```

## Evaluation proceduce

1. calculate the ratio between mae/rmse of reconciled forecasts and mae/rmse of base forecasts
2. mae/rmse of the total, mean, max and min mae/rmse of bottom series
3. summarised over specific domains

## mint and wlsv vs ols and wlss

```{r, include=FALSE}
store_base <- store_all %>% filter(rf_method == "base", accuracy_method == "mae") %>%
  rename(total_base = total, bottom_base = bottom) %>%
  select(batch, total_base, bottom_base)

mae_ratio <- store_all %>% filter(accuracy_method == "mae") %>%
  left_join(store_base, by = "batch") %>% 
  rowwise() %>%
  mutate(total = total/total_base, bottom = list(c(bottom/bottom_base))) %>%
  mutate(bottom_mean = mean(bottom), bottom_max = max(bottom),  bottom_min = min(bottom)) %>%
  ungroup()

store_base_rmse <- store_all %>% filter(rf_method == "base", accuracy_method == "rmse") %>%
  rename(total_base = total, bottom_base = bottom) %>%
  select(batch, total_base, bottom_base)

rmse_ratio <- store_all %>% filter(accuracy_method == "rmse") %>%
  left_join(store_base_rmse, by = "batch") %>% 
  rowwise() %>%
  mutate(total = total/total_base, bottom = list(c(bottom/bottom_base))) %>%
  mutate(bottom_mean = mean(bottom), bottom_max = max(bottom),  bottom_min = min(bottom)) %>%
  ungroup()
```

### OLS and structural scaling are more likely to produce more extreme results.

```{r}
mae_ratio %>% group_by(rf_method) %>% summarise(total= mean(total, na.rm=TRUE), bottom_max=mean(bottom_max, na.rm=TRUE), 
                                                bottom_min=mean(bottom_min, na.rm=TRUE), bottom_mean=mean(bottom_mean, na.rm=TRUE))
```
```{r}
rmse_ratio %>% group_by(rf_method) %>% summarise(total= mean(total, na.rm=TRUE), bottom_max=mean(bottom_max, na.rm=TRUE), 
                                                bottom_min=mean(bottom_min, na.rm=TRUE), bottom_mean=mean(bottom_mean, na.rm=TRUE))
```
### On average, mint and wlss don't improve the bottom level, but improve the total level.

### In general, the improvement of some series comes with the cost of some other series. It's very unlikely that reconciliation improves all the series.

```{r}
mae_ratio %>% filter(bottom_max <= 1.01, rf_method != "base") %>% select(representator, rf_method, distance, cluster, total, bottom_mean, bottom_max, bottom_min, batch) %>%
  arrange(total)
```
```{r}
rmse_ratio %>% filter(bottom_max <= 1.05, rf_method != "base") %>% select(representator, rf_method, distance, cluster, total, bottom_mean, bottom_max, bottom_min, batch) %>%
  arrange(total)
```

```{r}
rmse_ratio %>% filter(bottom_max <= 1.5, rf_method %in% c("ols", "wlss")) %>% select(representator, rf_method, distance, cluster, total, bottom_mean, bottom_max, bottom_min, batch) %>%
  arrange(total)
```

### robustness: wlsv > mint > wlss > ols

## K-medoids (only mint)

```{r}
rmse_ratio <- rmse_ratio %>% filter(rf_method == "mint") %>% select(-rf_method, -accuracy_method)
mae_ratio <- mae_ratio %>% filter(rf_method == "mint") %>% select(-rf_method, -accuracy_method)
```

```{r}
tmp <- rmse_ratio %>% filter(startsWith(cluster, "kmedoids")) %>%
  rowwise() %>% 
  mutate(n_middle = NROW(other$S)) %>%
  select(representator, distance, cluster, total, bottom_mean, bottom_max, bottom_min, n_middle, batch) %>%
  ungroup()
```

### distance:

```{r}
tmp %>% group_by(distance) %>% summarise_at(c("total", "bottom_max", "bottom_mean", "bottom_min"), c(mean=function(x){mean(x, na.rm=TRUE)}))
```

### representator: error > ts

```{r}
rmse_ratio %>% group_by(representator) %>% summarise_at(c("total", "bottom_max", "bottom_mean", "bottom_min"), c(mean=function(x){mean(x, na.rm=TRUE)}, min=function(x){min(x, na.rm=TRUE)}, max = function(x){max(x, na.rm=TRUE)}))
```

```{r}
tmp %>% group_by(n_middle) %>% summarise(total=mean(total, na.rm=TRUE), bottom_mean = mean(bottom_mean, na.rm=TRUE), bottom_max = mean(bottom_max, na.rm=TRUE), bottom_min = mean(bottom_mean, na.rm=TRUE))
```

```{r}

rmse_ratio %>% filter(cluster == "natural", batch == 0) %>% group_by(cluster)  %>% summarise(total=mean(total, na.rm=TRUE), bottom_mean = mean(bottom_mean, na.rm=TRUE), bottom_max = mean(bottom_max, na.rm=TRUE), bottom_min = mean(bottom_min, na.rm=TRUE))

```

```{r}
rmse_ratio %>% filter(cluster == "random-single-natural") %>% arrange(total, bottom_mean) %>% filter(batch == 0)
```