### 2023-09-26 analysis.Rmd
add tourism

---
title: "Constructing hierarchies"
author: "Bohan Zhang"
date: "2023-09-26"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

# Illustrative example

Consider a hierarchy with $4$ time series A,C,D,E, $A=C+D+E$. We consider
two kind of aggregation $B_1 = D+E$, $B_2 = C+D$.

```{r, include=FALSE, echo=FALSE}
S <- rbind(c(1,1,1), c(0, 1, 1), diag(3))

weights <- function(W) {
  W_inv <- solve(W)
  S %*% solve(t(S) %*% W_inv %*% S, t(S) %*% W_inv)
}

weights_plot <- function(offdiag, title = "Setting 1") {

  construct_W <- function(offdiag, x) {
    m <- matrix(0, 5, 5)
    m[lower.tri(m)] <- offdiag
    m[upper.tri(m)] <- t(m)[upper.tri(t(m))]
    diag(m) <- c(1, 1, 1, 1, 1)
    m[4, 5] <- x
    m[5, 4] <- x
    diag(sqrt(c(3, 2.1, 0.9, 1.05, 1.04))) %*% m %*% diag(sqrt(c(3, 2.1, 0.9, 1.05, 1.04)))
  }
  print(construct_W(offdiag, 0.5))

  weights_all <- vector("list", 5)

  for (i in seq(-0.99, 0.99, by = 0.01)) {
    W <- construct_W(offdiag, i)
    weight_i <- weights(W)
    for (j in 1:5){
      weights_all[[j]] <- rbind(weights_all[[j]], c(i, weight_i[,j]))
    }
  }
  library(dplyr)
  for (j in 1:5) {
    colnames(weights_all[[j]]) <- c("cor", "A", "B", "C", "D", "E")
    weights_all[[j]] <- as_tibble(weights_all[[j]]) %>%
      mutate(weights_col = c("A", "B", "C", "D", "E")[j])
  }

  weights_all <- do.call(rbind, weights_all)

  weights_all <- weights_all %>% tidyr::pivot_longer(names_to = "weights_row", cols = c("A", "B", "C", "D", "E"))

  library(ggplot2)

  ggplot(weights_all)  +
    geom_line(mapping = aes(x = cor, y = value, color = weights_row, group=weights_row)) +
    facet_wrap(~ weights_col) +
    ggtitle(title)
}
```

## Trend of weights

Consider the hierarchy $A, B_1, C, D, E$, $B_1 = D+E$ and variance $[3, 2.1, 0.9, 1.05, 1.04]'$

### Trend of weights: Scenario 1

All the correlation coefficients are $0$.

$$
\begin{bmatrix}
1 & 0 & 0 & 0 & 0 \\
0 &1 & 0 & 0 & 0 \\
0 & 0 & 1 & 0 & 0 \\
0 & 0 & 0 & 1 & x \\
0 & 0 & 0 & x & 1
\end{bmatrix}
$$

The trends of weights:

```{r, echo=FALSE, message=FALSE, results='hide'}
weights_plot(0)
```

### Trend of weights: Scenario 2

All the correlation coefficients are uniformly drawn from $[0, 0.5]$.

```{r, echo=FALSE, message=FALSE}
offdiag <- runif(10, 0, 0.5)
weights_plot(offdiag, title = "Setting 2")
```

### Trend of weights: Scenario 3

All the correlation coefficients are uniformly drawn from $[-0.5, 0]$.

```{r, echo=FALSE, message=FALSE}
offdiag <- runif(10, -0.5, 0)
weights_plot(offdiag, title = "Setting 3")
```

### Trend of weights: Scenario 4

All the correlation coefficients are uniformly drawn from $[-0.5, 0.5]$.

```{r, echo=FALSE, message=FALSE}
offdiag <- runif(10, -0.5, 0.5)
weights_plot(offdiag, title = "Setting 4")
```

### Conclusion

Overall, the weights for Series $B_1, D, E, A$ from Series $B_1$ ($G_{B_1,B_1}, G_{D,B_1}, G_{E,B_1}, G_{A, B_1}$) increase as the correlation between $D$ and $E$ increase. (See the panel B in each figure)

## forecast accuracy of two different grouping method: B1 and B2

```{r, echo=FALSE, message=FALSE}
source("R/reconciliation.R")
library(forecast)
step2 <- function(Wb) {

  errors <- MASS::mvrnorm(1000, rep(0, 3), Wb)

  AR <- diag(c(0, 0.3, 0.4))
  series <- matrix(0, 1001, 3)
  for (i in 2:1001) {
    series[i,] <- AR %*% series[i-1, ] + errors[i-1,]
  }
  series <- series[102:1001,]
  series[,1] <- seq(from=1, to=5, length = 900) + series[,1]
  series[,2] <- seq(from=1, to=5, length = 900) + series[,2]

  plot_series <- ts(series)
  colnames(series) <- c("C", "D", "E")
  plot(plot_series)

  S2 <- S
  S2[2,] <- c(1, 1, 0)

  series2 <- series %*% t(S2)
  series <- series %*% t(S)
  # Cross Validation
  init_window <- 800
  step <- 10

  mat2list <- function(x){
    r <- lapply(seq_len(ncol(x)), function(i) unname(x[,i]))
    names(r) <- colnames(x)
    r
  } 
  accs <- vector("list", 3)
  names(accs) <- c("base", "mint", "mint2")
  for (i in 1:10) {
    train <- series[1:(790+10*i),]
    test <- series[(791+10*i):(800+10*i),]

    fcasts <- lapply(iterators::iter(train, by="col"), function(s){
      mdl <- forecast(auto.arima(s, seasonal = FALSE), h=10)
      list(resid = as.numeric(residuals(mdl)), fcasts = as.numeric(mdl$mean))
    })

    resids <- do.call(cbind, lapply(fcasts, function(x){
      x$resid
    }))
    fcasts <- do.call(cbind, lapply(fcasts, function(x){
      x$fcasts
    }))

    fcasts_B2 <- forecast(auto.arima(series2[1:(790+10*i), 2], seasonal = FALSE), h=10)
    resids2 <- resids
    resids2[,2] <- as.numeric(residuals(fcasts_B2))
    fcasts2 <- fcasts
    fcasts2[,2] <- as.numeric(fcasts_B2$mean)

    reconf <- reconcile.mint(S, fcasts, resids)
    reconf2 <- reconcile.mint(S2, fcasts2, resids2)

    rmse <- function(x, y) { 
      r <- sapply(c(1, 3, 4, 5), function(i){
        sqrt(mean((x[,i]-y[,i])^2))
      })
      names(r) <- c("A", "C", "D", "E")
      r
    }
    accs$base <- rbind(accs$base, rmse(fcasts, test))
    accs$mint <- rbind(accs$mint, rmse(reconf, test))
    accs$mint2 <- rbind(accs$mint2, rmse(reconf2, test))
  }
  accs$base <- mat2list(accs$base)
  accs$mint <- mat2list(accs$mint)
  accs$mint2 <- mat2list(accs$mint2)
  as_tibble(accs) %>% mutate(names = c("A", "C", "D", "E")) %>% tidyr::unnest(cols = c("base", "mint", "mint2"))
}
```

Simulation procedure:

1. produce time series
2. test reconciliation accuracy for $B_1$ hierarchy and $B_2$ hierarchy through CV.

$$
\begin{bmatrix}
y_{C,t} \\ y_{D,t} \\ y_{E,t}
\end{bmatrix} = \begin{bmatrix}0.0055 \\ 0.0055 \\ 0\end{bmatrix}t + \begin{bmatrix}
0 & 0 & 0 \\
0 & 0.3 & 0 \\
0 & 0 & 0.4
\end{bmatrix} \begin{bmatrix}
y_{C,t-1} \\ y_{D,t-1} \\ y_{E,t-1}
\end{bmatrix} + \begin{bmatrix}
e_{C,t} \\ e_{D,t} \\ e_{E,t}
\end{bmatrix}
$$

$\boldsymbol{e}_t$ has the following covariance matrix

$$
\begin{bmatrix}
1 & 0 & 0 \\
0 & 1 & x \\
0 & x & 1 \\
\end{bmatrix}
$$

### x=0.9

```{r, message=FALSE, echo=FALSE}
W <- diag(3)
W[2,3] <- 0.9
W[3,2] <- 0.9
r1 <- step2(W)
r1 %>% group_by(names) %>% summarise_all(mean)
```

### x=-0.9

```{r, message=FALSE, echo=FALSE}
W <- diag(3)
W[2,3] <- -0.9
W[3,2] <- -0.9
r1 <- step2(W)
r1 %>% group_by(names) %>% summarise_all(mean)
```

## Mortality

```{r, echo=FALSE, message=FALSE}

library(dplyr)
library(tidyr)

# ts and error representator
source("R/construct_hierarchy.R", chdir = TRUE)

rf_method <- "mint"
accuracy_method <- "rmse"

# read all the results
store_all <- NULL
for (batch in 0:11) {
  store_all <- readRDS(paste0("mortality/store_", batch, ".rds"))$output %>%
    output_pre() %>% mutate(batch = batch) %>%
    filter(rf_method == .env$rf_method, accuracy_method == .env$accuracy_method) %>%
    dplyr::select(-rf_method, -accuracy_method) %>%
    rbind(store_all)
}

store_all <- store_all %>% rowwise() %>%
  mutate(values = list(c(total, bottom))) %>% ungroup() %>%
  dplyr::select(-total, -bottom)

## remove single random
store_all <- store_all %>% filter(!startsWith(cluster, "random-single"))

# store_all %>% group_by(representator, distance, cluster) %>% tally() %>%
#   pull(n) %>% table()

## combine the results of 12 batches for one clustering method

store_all <- store_all %>% select(-other) %>% nest(values = c(batch, values)) %>%
  mutate_at("values", purrr::map, function(x){
    x <- arrange(x, batch)
    do.call(c, x$values)
  })

## convert the tibble to a data.frame used for comparing ranks
store_mat <- do.call(cbind, store_all$values)
store_rank <- apply(store_mat, 1, rank) %>% t() %>% colMeans()
store_all$rank <- store_rank

## MCB test
mcb <- function(x) {
  tsutils::nemenyi(x, plottype = "vmcb")
}

method_desc <- function(row){
  print(sprintf("representator: %s, distance: %s, cluster: %s", 
                store_all$representator[row],
                store_all$distance[row],
                store_all$cluster[row]))
}
```

### Overall rank

```{r}
mcb(store_mat)
```

```{r}
store_all %>% arrange(rank)
```

### representator rank

```{r}
# representator rank
representator_summarise <- store_all %>% filter(distance != "") %>%
  nest(values = c(representator, values, rank)) %>%
  mutate_at("values", purrr::map, function(x){
    x$rank <- do.call(cbind, x$values) %>% apply(1, rank) %>%
      t() %>% colMeans()
    x %>% select(-values) %>%
      pivot_wider(names_from = "representator", values_from = "rank")
  }) %>%
  unnest("values")

representators <- c("error", "ts", "error.features", "ts.features", "forecast")

## in most cases, error is the best
representator_summarise %>% rowwise() %>%
  mutate(min = representators[which.min(c(error, ts, error.features, ts.features, forecast))]) %>%
  pull(min) %>%
  table()

## only in one case, error is the worst
representator_summarise %>% rowwise() %>%
  mutate(max = representators[which.max(c(error, ts, error.features, ts.features, forecast))]) %>%
  pull(max) %>%
  table()
mcb(representator_summarise %>% select(representators))

# no hierarchy is not good: 391 / 428
rank(store_all$rank)[1]

# natural hierarchy is not bad: 72/428
rank(store_all$rank)[2]

# random versus random natural versus natural

tmp <- store_all %>% filter(distance == "")
tmp$rank <- do.call(cbind, tmp$values) %>% apply(1, rank) %>%
  t() %>% colMeans()

## 1. all the averaged random hierarchies beat no hierarchy
## 2. deep and wide hierarchy wins
## 3. average of more hierarchies does not obtain better results. 
##   Some really bad hierarchies destroy the results. 
##.  Deeper hierarchies have the ability to mitigate the affect of bad series

```

### cluster rank
```{r}
# cluster rank
cluster_summarise <- store_all %>% filter(distance != "") %>%
  nest(values = c(cluster, values, rank)) %>%
  mutate_at("values", purrr::map, function(x){
    x$rank <- do.call(cbind, x$values) %>% apply(1, rank) %>%
      t() %>% colMeans() %>% rank()
    x %>% select(-values) %>%
      pivot_wider(names_from = "cluster", values_from = "rank")
  }) %>%
  unnest("values")

cluster_summarise %>% select(-c(representator, distance)) %>%
  mcb()

## 1. kmedoids unnested is really bad
## 2. hcluster with single linkage is bad
## 3. nested kmedoids performs really well, which forms a natural-like unbalanced hierarchy.
```

### distance rank

```{r}
distance_summarise <- store_all %>% filter(distance != "") %>%
  nest(values = c(distance, values, rank)) %>%
  mutate_at("values", purrr::map, function(x){
    x$rank <- do.call(cbind, x$values) %>% apply(1, rank) %>%
      t() %>% colMeans() %>% rank()
    x %>% select(-values) %>%
      pivot_wider(names_from = "distance", values_from = "rank")
  }) %>%
  unnest("values")
distance_summarise %>% select(-c(representator, cluster)) %>%
  mcb()

distances <- c("euclidean", "dtw", "negcor", "cor", "uncorrelation")
distance_summarise %>% rowwise() %>%
  mutate(max = distances[which.max(c(euclidean, dtw, negcor, cor, uncorrelation))]) %>%
  pull(max) %>%
  table()
distance_summarise %>% rowwise() %>%
  mutate(min = distances[which.min(c(euclidean, dtw, negcor, cor, uncorrelation))]) %>%
  pull(min) %>%
  table()
```

## Tourism

```{r, include=FALSE, echo=FALSE}
store_all <- NULL
for (batch in 0:8) {
  store_all <- readRDS(paste0("tourism/store_", batch, ".rds"))$output %>%
    output_pre() %>% mutate(batch = batch) %>%
    filter(rf_method == .env$rf_method, accuracy_method == .env$accuracy_method) %>%
    dplyr::select(-rf_method, -accuracy_method) %>%
    rbind(store_all)
}

store_all <- store_all %>% rowwise() %>%
  mutate(values = list(c(total, bottom))) %>% ungroup() %>%
  dplyr::select(-total, -bottom)

## remove single random
store_all <- store_all %>% filter(!startsWith(cluster, "random-single"))

# store_all %>% group_by(representator, distance, cluster) %>% tally() %>%
#   pull(n) %>% table()

## combine the results of 12 batches for one clustering method

store_all <- store_all %>% dplyr::select(-other) %>% nest(values = c(batch, values)) %>%
  mutate_at("values", purrr::map, function(x){
    x <- arrange(x, batch)
    do.call(c, x$values)
  })

## convert the tibble to a data.frame used for comparing ranks
store_mat <- do.call(cbind, store_all$values)
store_rank <- apply(store_mat, 1, rank) %>% t() %>% colMeans()
store_all$rank <- store_rank
```

```{r}
mcb(store_mat)
```
```{r}
store_all %>% arrange(rank)
```

### representator rank

```{r, message=FALSE}
# representator rank
representator_summarise <- store_all %>% filter(distance != "") %>%
  nest(values = c(representator, values, rank)) %>%
  mutate_at("values", purrr::map, function(x){
    x$rank <- do.call(cbind, x$values) %>% apply(1, rank) %>%
      t() %>% colMeans()
    x %>% dplyr::select(-values) %>%
      pivot_wider(names_from = "representator", values_from = "rank")
  }) %>%
  unnest("values")

representators <- c("error", "ts", "error.features", "ts.features", "forecast")

## in most cases, error is the best
representator_summarise %>% rowwise() %>%
  mutate(min = representators[which.min(c(error, ts, error.features, ts.features, forecast))]) %>%
  pull(min) %>%
  table()

## only in one case, error is the worst
representator_summarise %>% rowwise() %>%
  mutate(max = representators[which.max(c(error, ts, error.features, ts.features, forecast))]) %>%
  pull(max) %>%
  table()
mcb(representator_summarise %>% dplyr::select(all_of(representators)))

# no hierarchy
rank(store_all$rank)[which(store_all$cluster == "")]

# natural hierarchy
rank(store_all$rank)[which(store_all$cluster == "natural")]

# random versus random natural versus natural

tmp <- store_all %>% filter(distance == "")
tmp$rank <- do.call(cbind, tmp$values) %>% apply(1, rank) %>%
  t() %>% colMeans()

## 1. all the averaged random hierarchies beat no hierarchy
## 2. deep and wide hierarchy wins
## 3. average of more hierarchies does not obtain better results. 
##   Some really bad hierarchies destroy the results. 
##.  Deeper hierarchies have the ability to mitigate the affect of bad series

```

### cluster rank
```{r}
# cluster rank
cluster_summarise <- store_all %>% filter(distance != "") %>%
  nest(values = c(cluster, values, rank)) %>%
  mutate_at("values", purrr::map, function(x){
    x$rank <- do.call(cbind, x$values) %>% apply(1, rank) %>%
      t() %>% colMeans() %>% rank()
    x %>% dplyr::select(-values) %>%
      pivot_wider(names_from = "cluster", values_from = "rank")
  }) %>%
  unnest("values")

cluster_summarise %>% dplyr::select(-c(representator, distance)) %>%
  mcb()

## 1. kmedoids unnested is really bad
## 2. hcluster with single linkage is bad
## 3. nested kmedoids performs really well, which forms a natural-like unbalanced hierarchy.
```

### distance rank

```{r}
distance_summarise <- store_all %>% filter(distance != "") %>%
  nest(values = c(distance, values, rank)) %>%
  mutate_at("values", purrr::map, function(x){
    x$rank <- do.call(cbind, x$values) %>% apply(1, rank) %>%
      t() %>% colMeans() %>% rank()
    x %>% select(-values) %>%
      pivot_wider(names_from = "distance", values_from = "rank")
  }) %>%
  unnest("values")
distance_summarise %>% select(-c(representator, cluster)) %>%
  mcb()

distances <- c("euclidean", "dtw", "negcor", "cor", "uncorrelation")
distance_summarise %>% rowwise() %>%
  mutate(max = distances[which.max(c(euclidean, dtw, negcor, cor, uncorrelation))]) %>%
  pull(max) %>%
  table()
distance_summarise %>% rowwise() %>%
  mutate(min = distances[which.min(c(euclidean, dtw, negcor, cor, uncorrelation))]) %>%
  pull(min) %>%
  table()
```