### 2024-04-15 manuscript/titlepage.tex
submit to IJF

\documentclass[a4paper,review,12pt,authoryear]{elsarticle}

% \usepackage{natbib}
\usepackage{amsfonts,amsmath,bbm,bm,xcolor,booktabs,hyperref,amsthm}
\usepackage{geometry}
\usepackage{todonotes}
\usepackage{hyperref}
\usepackage{ulem}
\usepackage{subfig}
\geometry{a4paper,scale=0.8}
\usepackage{setspace}
\setstretch{1.5}
\let\code=\texttt
\let\proglang=\textsf
\setcounter{MaxMatrixCols}{20}
\usepackage{enumitem}
% \usepackage[pagewise]{lineno}

\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  citecolor=blue,
  urlcolor=blue}

\begin{document}

\begin{frontmatter}

  %\title{On the performance of hierarchy construction: impacts of structure and clustering}
\title{Constructing hierarchical time series through clustering: \\Is there an optimal way for forecasting?}

\author[label1]{Bohan Zhang\corref{cor1}}
\address[label1]{School of Economics and Management, Beihang University, Beijing, China}
\ead{zhangbohan@buaa.edu.cn}
\cortext[cor1]{Corresponding author.}
\author[label2]{Anastasios Panagiotelis}
\address[label2]{The University of Sydney Business School, NSW 2006, Australia}
\author[label3]{Han Li}
\address[label3]{Department of Economics, The University of Melbourne, VIC 3010, Australia}

\end{frontmatter}

\section*{Acknowledgements}

Bohan Zhang is supported by the international joint doctoral education fund of Beihang University. Han Li is supported by the Australian Research Council Discovery Projects funding scheme (DP220100090).

\section*{Declaration of competing interest}

The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.

\end{document}