### 2024-09-08 Revision1/coverletter.tex

\documentclass[a4paper,12pt]{amsart}
\usepackage{natbib}
\usepackage{graphics, graphicx}
\usepackage{epstopdf}
\usepackage{float}
\usepackage{bm}
\usepackage{amsmath,amsfonts}
\usepackage{geometry}
\usepackage{amsmath,amsfonts,amssymb}
%\usepackage{amsmath,amsfonts,amssymb,stmaryrd,comment,tabularx, caption, rotating}
\usepackage{setspace,a4,times}
%\usepackage{harvard}
\usepackage{graphicx}
\usepackage{url}
\usepackage{morefloats}
\usepackage{float}
\usepackage{subfigure}
 \setlength{\topmargin}{-0.5cm}
% \setlength{\topskip}{1cm}
% \setlength{\headheight}{0cm}
% \setlength{\headsep}{0cm}
 \setlength{\oddsidemargin}{-0.2 cm}
 \setlength{\evensidemargin}{-0.2 cm}
 \setlength{\textwidth}{16cm}
 \setlength{\textheight}{25cm}
  %\setlength{\textheight}{25cm}
%\renewcommand{\baselinestretch}{2.0}
\geometry{top=3cm,bottom=2.5cm, left=3.4cm,right=3.4cm}
\setlength{\parindent}{0pt}
\linespread{1.35}
\newcommand{\bx}{\bm{x}}
\newcommand{\bv}{\bm{v}}
\newcommand{\bg}{\bm{g}}
\newcommand{\bbeta}{\bm{\beta}}
\newcommand{\bz}{\bm{z}}
\usepackage{color}

\doublespacing

\begin{document}

\begin{flushright}
Mr Bohan Zhang\\
Beihang University\\
School of Economics and Management\\%, Department of Economics\\
Xueyuan Road 37, Haidian District \\
Beijing, China\\
5 July 2024
\end{flushright}

%Professor David Blake and Professor Andrew Cairns\\
%Co-Editor\\
%Insurance: Mathematics and Economics\\
\bigskip
RE: ``Constructing hierarchical time series through clustering: Is there an optimal way for forecasting?''   \\

Dear Professor Fotios Petropoulos,\\

Thank you for giving us the opportunity to revise and resubmit this paper. We hope that the incorporation of the suggested changes will now make the revised paper suitable for \textit{International Journal of Forecasting}. We have addressed the comments from the Associate Editor and the two reviewers in separate files.    \\

We greatly appreciate the time and detail provided by you, and we look forward to your response.\\

Yours sincerely,\\
Bohan Zhang

\end{document}