### 2024-09-08 submission/coverletter.tex

\documentclass[a4paper,12pt]{amsart}
\usepackage{natbib}
\usepackage{graphics, graphicx}
\usepackage{epstopdf}
\usepackage{float}
\usepackage{bm}
\usepackage{amsmath,amsfonts}
\usepackage{geometry}
\usepackage{amsmath,amsfonts,amssymb}
%\usepackage{amsmath,amsfonts,amssymb,stmaryrd,comment,tabularx, caption, rotating}
\usepackage{setspace,a4,times}
%\usepackage{harvard}
\usepackage{graphicx}
\usepackage{url}
\usepackage{morefloats}
\usepackage{float}
\usepackage{subfigure}
 \setlength{\topmargin}{-0.5cm}
% \setlength{\topskip}{1cm}
% \setlength{\headheight}{0cm}
% \setlength{\headsep}{0cm}
 \setlength{\oddsidemargin}{-0.2 cm}
 \setlength{\evensidemargin}{-0.2 cm}
 \setlength{\textwidth}{16cm}
 \setlength{\textheight}{25cm}
  %\setlength{\textheight}{25cm}
%\renewcommand{\baselinestretch}{2.0}
\geometry{top=3cm,bottom=2.5cm, left=3.4cm,right=3.4cm}
\setlength{\parindent}{0pt}
\linespread{1.35}
\newcommand{\bx}{\bm{x}}
\newcommand{\bv}{\bm{v}}
\newcommand{\bg}{\bm{g}}
\newcommand{\bbeta}{\bm{\beta}}
\newcommand{\bz}{\bm{z}}
\usepackage{color}

\doublespacing

\begin{document}

\begin{flushright}
Mr Bohan Zhang\\
Beihang University\\
School of Economics and Management\\%, Department of Economics\\
Xueyuan Road 37, Haidian District \\
Beijing, China\\
5 July 2024
\end{flushright}

%Professor David Blake and Professor Andrew Cairns\\
%Co-Editor\\
%Insurance: Mathematics and Economics\\
\bigskip
RE: ``Constructing hierarchical time series through clustering: Is there an optimal way for forecasting?''   \\

Dear Professor Fotios Petropoulos,\\

Thank you for conditionally accepting our paper for publication in \textit{International Journal of Forecasting}. We have made a reproducibility package available in an online GitHub repository, which can be accessed at \url{https://github.com/AngelPone/project_hierarchy}.    \\

We greatly appreciate the time and detail provided by you, and we look forward to your response.\\

Yours sincerely,\\
Bohan Zhang

\end{document}